---
title: "Exercises for data manipulation - E1"
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(webexercises)
```

### Exercise 1.1

Load the package `tidyr` (or go crazy and just install the whole tidyverse). Run the following piece of code which creates a data set called `persons`.

```{r class.source = 'fold-show', message=FALSE}

persons <- tribble(
  ~name,             ~measure,  ~value,
  "Anne Smith", "age", 45,
  "Anne Smith", "height in cm", 171,
  "Peter Rowe", "age", 37,
  "Peter Rowe", "height in cm", 166,
  "Virginia Castles", "age", 20,
  "Virginia Castles", "height in cm", 168
)
```

Make the data wider, so that each person is one row.

`r hide("Hint")`
Use `pivot_wider()`. Potentially look at what the argument `names_repair` does.
`r unhide()`

`r hide("Answer")`
```{r}

persons |> 
  pivot_wider(names_from = measure, values_from = value)

```

"height in cm" is not a standard variable name and can become cumbersome to work with. If this bothers us, we can use the argument `names_repair` for repairing the names.

```{r}
persons |> 
  pivot_wider(names_from = measure, values_from = value, names_repair = "universal")
```
`r unhide()`

### Exercise 1.2

Repeat the exercise with this table - here, several persons have the same name, and the measurements have an associated uncertainty.

```{r}
persons <- tribble(
  ~name, ~by, ~measure,  ~value, ~uncertainty,
  "Anne Smith", "London",   "age",       45, 0,
  "Anne Smith", "London",   "height in cm", 171, 0,
  "Anne Smith", "Leeds", "age",       88, 0,
  "Anne Smith", "Leeds", "height in cm", 154, 0.5,
  "Peter Rowe", "Exeter", "age",       37, 0,
  "Peter Rowe", "Exeter", "height in cm", 166, 1.3,
  "Virginia Castles", "Bristol",  "age",       20, 1,
  "Virginia Castles", "Bristol",  "height in cm", 168, 1.4
)
```

`r hide("Answer")`

```{r}
persons |> 
  pivot_wider(names_from = measure, values_from = c(value, uncertainty), names_repair = "universal")
```

Be aware that `pivot_wider` assumes, that anything not included in `values_from` or `names_from` is an ID variable. Therefore, it can tell the difference between Anne in London and Anne in Leeds.

If we want more control over the names of the variables we create, we can use the `names_glue` argument.
```{r}
persons |> 
  pivot_wider(names_from = measure, values_from = c(value, uncertainty), names_repair = "universal",
              names_glue = "{measure}_{.value}")
```
`r unhide()`

### Exercise 1.3a

The data set `relig_income` is built in to `tidyr`. The documentation can be seen with `?relig_income`. 

Transform the data set so that each row contains the number of respondents in each combination of income group and religion

```{r webex.hide="Answer"}
relig_income_long <- relig_income |> 
  pivot_longer(cols = -religion, names_to = "income", values_to = "respondents")
```

### Exercise 1.3b

Use the resulting table to calculate the number of respondents in each religious group, each income group, and the corresponding percentages.

`r hide("Hint")`
Use `group_by()` and `summarise()`
`r unhide()`

```{r eval=FALSE, webex.hide="Answer"}

relig_income_long |> 
  group_by(religion) |> 
  summarise(respondents = sum(respondents)) |>
  mutate(pct = 100 * respondents / sum(respondents))

relig_income_long |> 
  group_by(income) |> 
  summarise(respondents = sum(respondents))|>
  mutate(pct = 100 * respondents / sum(respondents))

```

### Exercise 1.3c

Create a cross table showing how many answered / didn't answer each of the two variables.

`r hide("Hint")`
Use mutate to create variables on whether the questions have been answered or not. Use group_by and summarise to sum up, then use pivot_wider on the results.
`r unhide()`

`r hide("Answer")`

```{r}

relig_income_long |> 
  mutate(religion = if_else(religion == "Don’t know/refused", "Didn't answer", "Answered"),
         income = if_else(income == "Don't know/refused", "Didn't answer", "Answered")) |> 
  group_by(religion, income) |> 
  summarise(respondents = sum(respondents)) |> 
  ungroup() |> 
  pivot_wider(values_from = respondents, names_from = income)

```
`r unhide()`